---
title: "pppgram-package: basic functionality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{basic_functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height=5,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggplot2)
library(pppgram)
library(spatstat)
```

Demonstrate the basic functionality of the package. Note that only 2D rectangular windows are supported at the moment.

Example point pattern:

```{r, fig.width=7, fig.height=4}
set.seed(1)
x <- rThomas(0.01, 3, 3, win = owin(c(0,200), c(0,100)))
plot(x)
```


# Estimation of 2D periodogram


## Bartlett's periodogram


```{r}
per0 <- periodogram(x, M = 15)
per0
```

If non-Fourier grid wavenumbers are used (or a non-rectangular window is used) the border-bias needs to be addressed

```{r, fig.width=6, fig.height=6}
per1 <- periodogram(x, omega = seq(-1, 1, l = 51)*.15, debias_early = FALSE) # will have bias near w=0
per2 <- periodogram(x, omega = per1$omega) # the default is TRUE
```

## Smoothing
This is simple smoother application using spatstat::Smooth.im (using spatstat::blur). A wrapper is provided:

```{r}
per2s <- Smooth(per2, kernel = "gaussian", sigma = 0.005)
```

## Multitapered periodogram

At the moment only sine-family of tapers is implemented. Choose number $P$ for $P^d$ tapers:

```{r}
per4 <- periodogram_mt(x, P = 2, omega = per1$omega)
per4
```


# Plotting 2D

Default plotting: The 2D periodograms are converted to spatstat::im objects and the plot.im is called.

```{r, fig.width=6, fig.height=6}
plot(spatstat::listof(basic=per0, biased=per1, debiased = per2, 
                      smoothed=per2s, multitapered=per4), main = "")
```

Does not provide axes-ticks. Alternatively, one can e.g. use ggplot like so:

```{r, fig.width=8, fig.height=3}
d1 <- rbind(
  as.data.frame(per2),
  as.data.frame(per4)
  )
  

ggplot(d1, aes(w1, w2, fill = sdf_estimate)) + geom_raster()  + coord_fixed(expand = FALSE) +
  facet_wrap(~type)
```

# Estimation of periodogram in wavenumber amplitude only (isotropic/rotationally averaged)

## Rotational averaging

We can simply average the 2D periodogram rotationally. 

```{r}
per2i <- rotmean.pppgram(per2, adjust = 2) # divides by lambda by default
# pretty much same as spatstat's
per2j <- rotmean(as.im(per2) / intensity(x)) # divide this also by lambda

plot(per2i)
lines(per2j, col = 3)
```

## Directly using isotropy assumption

Default is to use a squared-exponential taper (set `taper_a=0` for no taper). The current implementation is a bit slow so might prefer the 2D+rotmean approach.

```{r}
peri <- periodogram_iso(x,  taper_a = 25)
peri0 <- periodogram_iso(x,  taper_a = 0)
peri
```

The results are spatstat::fv - class, so plotting e.g.

```{r}
plot(peri)
lines(peri0, col = 3)
lines(per2i, col = 4, lty = 2)
```

