---
title: "Basic functionality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height=5,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(ggplot2)
library(pppgram)
library(spatstat)
```

Demonstrate the basic functionality of the package. Note that only 2D rectangular windows are supported at the moment.

Example point pattern:

```{r, fig.width=7, fig.height=4}
set.seed(1)
x <- rThomas(0.01, 3, 3, win = owin(c(0,200), c(0,100)))
plot(x)
```


# Estimation of 2D periodogram


## Bartlett's periodogram


```{r}
per0 <- periodogram(x, M = 25)
per0
```

If non-Fourier grid wavenumbers are used (or a non-rectangular window is used) the border-bias needs to be addressed

```{r, fig.width=6, fig.height=6}
per1 <- periodogram(x, omega = seq(-1, 1, l = 51)*.15, debias_early = FALSE) # will have bias near w=0
per2 <- periodogram(x, omega = per1$omega) # the debias_early default is TRUE
```

## Smoothing
This is simple smoother application using `spatstat.explore::Smooth.im`. A wrapper is provided:

```{r}
per2s <- Smooth(per2, kernel = "gaussian", sigma = 0.005)
```

## Multitapered periodogram

At the moment only sine-family of tapers is implemented. Choose number $P$ for $P^d$ tapers:

```{r}
per4 <- periodogram_mt(x, P = 2, omega = per1$omega)
per4
```


# Plotting 2D

Default plotting: The 2D periodograms are converted to spatstat::im objects and the plot.im is called.

```{r, fig.width=6, fig.height=4}
plot(spatstat.geom::listof(basic=per0, biased=per1, debiased = per2, 
                      smoothed=per2s, multitapered=per4), main = "")
```

Does not provide axes-ticks. Alternatively, one can use `as.data.frame` method and plot with e.g. `ggplot`:

```{r, fig.width=8, fig.height=3}
d1 <- rbind(
  as.data.frame(per2),
  as.data.frame(per4)
  )
  
ggplot(d1, aes(w1, w2, 
               fill = sdf_estimate )) + 
  geom_raster()  + 
  coord_fixed(expand = FALSE) +
  scale_fill_viridis_c(option = "A") +
  facet_wrap(~type)
```

# Estimation of periodogram in wavenumber amplitude only (isotropic/rotationally averaged)

## Rotational averaging

A reasonable approach is to average the 2D periodogram rotationally. 

```{r}
per2i <- rotmean.pppgram(per2, adjust = 1) # divides by lambda by default
# pretty much same as spatstat's, bw arbitrary here
per2j <- rotmean(as.im(per2) / intensity(x), adjust = 5) # divide this also by lambda

plot(per2i)
lines(per2j, col = 3)
```

## Directly using isotropy assumption

Default is to use a squared-exponential taper (set `taper_a=0` for no taper). The current implementation is a bit slow so might prefer the 2D+rotmean approach.

```{r}
peri <- periodogram_iso(x,  taper_a = 25)
peri0 <- periodogram_iso(x,  taper_a = 0)
peri
```

The results are in `spatstat::fv` - class, so plotting e.g.

```{r}
plot(peri)
lines(peri0, col = 3)
lines(per2i, col = 4, lty = 2)
```

# Ring-pass filtering

The sdf at low wavenumbers captures the clustering at higher spatial scales. 

We can visualise the data pattern's "signal" at those wavenumbers using a filter.

```{r, fig.width=4, fig.height=3}
lo <- ringpass(x, a = 0, b = 0.05, nx = 2^9)
# for comparison, higher wavenumber filter
hi <- ringpass(x, a = 0.15, b= 0.2, nx = 2^9)

co <- hcl.colors(12, "Mint")  

par(mfrow= c(2,1), mar = c(0,0,1,0))
plot(lo, col = co, ribbon = FALSE)
points(x, cex = .05)
plot(hi, col = co, ribbon = FALSE)
```




